### 2024-09-24 outputs/paper/paper.qmd
cleaned folder structure, write final cleaned and combined data set

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Robert Ford
thanks: "Code and data are available at: https://github.com/Ford-Robert/STA304_City-Of-Toronto-Data.git"
date: today
date-format: long??
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

# Data {#sec-data}

# Discussion {#sec-discussion}

## First discussion point {#sec-first-point}

## Second discussion point

## Third discussion point

## Weaknesses and next steps

\newpage

# References