### 2024-09-25 outputs/paper/paper.qmd
added some graphs

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Robert Ford
thanks: "Code and data are available at: https://github.com/Ford-Robert/STA304_City-Of-Toronto-Data.git"
date: today
date-format: long??
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(lubridate)
library(opendatatoronto)
library(tidyverse)
library(tidyr)
library(dplyr)
library(knitr)
library(ggplot2)
library(scales)
library(gridExtra)
library(patchwork)
```

```{r}
data <- read_csv(here::here("outputs/data/cleaned_dataset.csv"))

```

# Introduction

# Data {#sec-data}

```{r num-incidents-by-mode}
plot1 <- ggplot(data, aes(x = vehicle)) +
  geom_bar(fill = "steelblue") +
  scale_y_continuous(labels = comma) +
  labs(title = "Number of Incidents by Vehicle Type", x = "Vehicle Type", y = "Number of Incidents") +
  theme_minimal()

cumulative_delay <- data %>%
  group_by(vehicle) %>%
  summarise(total_delay = sum(Delay, na.rm = TRUE))

plot2 <- ggplot(cumulative_delay, aes(x = vehicle, y = total_delay)) +
  geom_bar(stat = "identity", fill = "tomato") +
  scale_y_continuous(labels = comma) +
  labs(title = "Total Cumulative Delay Time by Vehicle", x = "Vehicle Type", y = "Total Delay (Minutes)") +
  theme_minimal()

grid.arrange(plot1, plot2, ncol = 2)

```

```{r ave-delay-by-mode}

average_delay <- data %>%
  group_by(vehicle) %>%
  summarise(avg_delay = mean(Delay, na.rm = TRUE))

ggplot(average_delay, aes(x = vehicle, y = avg_delay)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Average Delay Time by Vehicle Type", x = "Vehicle Type", y = "Average Delay (Minutes)") +
  theme_minimal()

```

```{r ave-delay-by-incident}
data_incident_avg_delay <- data %>%
  group_by(Incident) %>%
  summarise(avg_delay = mean(Delay, na.rm = TRUE), .groups = 'drop') %>%
  mutate(Incident = fct_reorder(Incident, avg_delay, .desc = TRUE))  # Reorder by avg_delay

# Create the bar chart
ggplot(data_incident_avg_delay, aes(x = Incident, y = avg_delay)) +
  geom_bar(stat = "identity", fill = "steelblue") +  # Set a single color for all bars
  labs(title = "Average Delay Time by Incident Type", 
       x = "Incident Type", y = "Average Delay (Minutes)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10))

data_incident_by_delay <- data %>%
  group_by(Incident, vehicle) %>%
  summarise(incident_count = n(), avg_delay = mean(Delay, na.rm = TRUE), .groups = 'drop') %>%
  group_by(Incident) %>%
  mutate(percentage = (incident_count / sum(incident_count)) * 100) %>%
  mutate(Incident = fct_reorder(Incident, avg_delay, .desc = TRUE))  # Reorder by avg_delay

# Create the stacked bar chart with percentages
ggplot(data_incident_by_delay, aes(x = Incident, y = avg_delay, fill = vehicle)) +
  geom_bar(stat = "identity", position = "fill") +  # Stacked bars, normalized to show proportions
  labs(title = "Average Delay Time by Incident Type and Vehicle (Proportions)",
       x = "Incident Type", y = "Proportion of Incidents", fill = "Vehicle Type") +
  scale_y_continuous(labels = scales::percent) +  # Show percentages on y-axis
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10))

```
```{r}
data_total_delay <- data %>%
  group_by(Incident, vehicle) %>%
  summarise(total_delay = sum(Delay, na.rm = TRUE), .groups = 'drop')

# Step 2: Reorder the Incident types based on total delay for better visualization
data_total_delay <- data_total_delay %>%
  group_by(Incident) %>%
  mutate(Incident = fct_reorder(Incident, total_delay, .desc = TRUE))  # Reorder by total_delay

# Step 3: Create the stacked bar chart
ggplot(data_total_delay, aes(x = Incident, y = total_delay, fill = vehicle)) +
  geom_bar(stat = "identity", position = "stack") +  # Stacked bars for total delay time
  labs(title = "Total Delay Time by Incident Type and Vehicle", 
       x = "Incident Type", y = "Total Delay Time (Minutes)", fill = "Vehicle Type") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10))
```

```{r incident-type}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: Delay Incidence by Month
#| fig-align: center
# Prepare the data: calculate the percentage of incidents by type for each vehicle
data_incident_percentage <- data %>%
  group_by(vehicle, Incident) %>%
  summarise(incident_count = n(), .groups = 'drop') %>%
  group_by(vehicle) %>%
  mutate(percentage = (incident_count / sum(incident_count)) * 100)

# Lump incidents making up less than 5% into "Other"
data_incident_percentage <- data_incident_percentage %>%
  mutate(Incident = ifelse(percentage < 5, "Other", Incident)) %>%
  group_by(vehicle, Incident) %>%
  summarise(incident_count = sum(incident_count), percentage = sum(percentage), .groups = 'drop')

# Function to create doughnut chart for a specific vehicle
create_doughnut_chart <- function(vehicle_type) {
  # Filter data for the specific vehicle type
  vehicle_data <- data_incident_percentage %>%
    filter(vehicle == vehicle_type)

  # Create the doughnut chart
  ggplot(vehicle_data, aes(x = 2, y = percentage, fill = Incident)) +
    geom_bar(stat = "identity", width = 1) +
    coord_polar(theta = "y") +
    xlim(0.5, 2.5) +  # Adjust the x-axis limits to create the doughnut shape
    geom_text(aes(label = paste0(Incident, "\n", round(percentage, 1), "%")), 
              position = position_stack(vjust = 0.5), size = 3) +
    labs(title = paste("Incident Distribution for", vehicle_type)) +
    theme_void() +  # Clean theme for the doughnut chart
    theme(legend.position = "none")  # Hide the legend for a cleaner look
}

# Create doughnut charts for each vehicle type
bus_doughnut <- create_doughnut_chart("Bus")
streetcar_doughnut <- create_doughnut_chart("Streetcar")
subway_doughnut <- create_doughnut_chart("Subway")

# Arrange the charts side by side
bus_doughnut + streetcar_doughnut + subway_doughnut

```

```{r ave-time-delay}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: Delay Incidence by Month
#| fig-align: center
# Prepare the data: calculate the average delay time for each incident type and vehicle
data_incident_avg_delay <- data %>%
  group_by(vehicle, Incident) %>%
  summarise(avg_delay = mean(Delay, na.rm = TRUE), .groups = 'drop')

# Function to create a bar chart for a specific vehicle
create_bar_chart <- function(vehicle_type) {
  # Filter data for the specific vehicle type and reorder incidents by average delay
  vehicle_data <- data_incident_avg_delay %>%
    filter(vehicle == vehicle_type) %>%
    mutate(Incident = fct_reorder(Incident, avg_delay, .desc = TRUE))  # Reorder by avg_delay

  # Create the bar chart with uniform color and spread the x-axis labels
  ggplot(vehicle_data, aes(x = Incident, y = avg_delay)) +
    geom_bar(stat = "identity", fill = "steelblue") +  # Set a single color for all bars
    labs(title = paste("Average Delay by Incident Type for", vehicle_type),
         x = "Incident Type", y = "Average Delay (Minutes)") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10))  # Rotate labels for readability
}

# Generate the separate bar charts for each vehicle type
bus_bar_chart <- create_bar_chart("Bus")
streetcar_bar_chart <- create_bar_chart("Streetcar")
subway_bar_chart <- create_bar_chart("Subway")

# Print each graph individually
print(bus_bar_chart)
print(streetcar_bar_chart)
print(subway_bar_chart)

```

```{r incidents-over-time}
data_inc_num <- data %>%
  group_by(month = floor_date(Date, "month"), vehicle) %>%
  summarise(incident_count = n())

# Create the line graph
ggplot(data_inc_num, aes(x = month, y = incident_count, color = vehicle, group = vehicle)) +
  geom_line(size = 1) +
  labs(title = "Number of Incidents Over Time by Vehicle", x = "Time", y = "Number of Incidents", color = "Vehicle Type") +
  theme_minimal()

data_monthly_delay <- data %>%
  group_by(month = floor_date(Date, "month"), vehicle) %>%
  summarise(total_delay = sum(Delay, na.rm = TRUE), .groups = 'drop')

# Create the line graph for monthly total delay times
ggplot(data_monthly_delay, aes(x = month, y = total_delay, color = vehicle, group = vehicle)) +
  geom_line(size = 1) +
  labs(title = "Total Delay Time Over Time by Vehicle (Monthly)", 
       x = "Time (Month)", y = "Total Delay Time (Minutes)", color = "Vehicle Type") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
data_total_delay <- data %>%
  group_by(month = floor_date(Date, "month")) %>%  # Group by month
  summarise(total_delay = sum(Delay, na.rm = TRUE), .groups = 'drop')

# Create the line chart
ggplot(data_total_delay, aes(x = month, y = total_delay)) +
  geom_line(size = 1, color = "steelblue") +  # Single line with color
  labs(title = "Total Delay Time Over Time", 
       x = "Time (Month)", y = "Total Delay Time (Minutes)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# Discussion {#sec-discussion}

## First discussion point {#sec-first-point}

## Second discussion point

## Third discussion point

## Weaknesses and next steps

\newpage

# References