### 2024-09-26 outputs/paper/paper.qmd
added graphs

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Robert Ford
thanks: "Code and data are available at: https://github.com/Ford-Robert/STA304_City-Of-Toronto-Data.git"
date: today
date-format: long??
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(lubridate)
library(opendatatoronto)
library(tidyverse)
library(tidyr)
library(dplyr)
library(knitr)
library(ggplot2)
library(scales)
library(gridExtra)
library(patchwork)
```

```{r}
#| message: false
#| echo: false
#| warning: false
data <- read_csv(here::here("outputs/data/cleaned_dataset.csv"))

```

# Introduction{#sec-introduction}

# Data {#sec-data}

```{r num-incidents-by-mode}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: (left) Total Number of Incidents (right) Total Amount of Delay Time
#| fig-align: center

plot1 <- ggplot(data, aes(x = vehicle)) +
  geom_bar(fill = "steelblue") +
  scale_y_continuous(labels = comma) +
  labs(title = "Number of Incidents by Vehicle Type", x = "Vehicle Type", y = "Number of Incidents") +
  theme_minimal()

cumulative_delay <- data %>%
  group_by(vehicle) %>%
  summarise(total_delay = sum(Delay, na.rm = TRUE))

plot2 <- ggplot(cumulative_delay, aes(x = vehicle, y = total_delay)) +
  geom_bar(stat = "identity", fill = "tomato") +
  scale_y_continuous(labels = comma) +
  labs(title = "Cumulative Delay Time by Vehicle", x = "Vehicle Type", y = "Total Delay (Minutes)") +
  theme_minimal()

grid.arrange(plot1, plot2, ncol = 2)

```

```{r ave-delay-by-mode}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: Average Amount of Delay Time by Vehicle Type
#| fig-align: center

average_delay <- data %>%
  group_by(vehicle) %>%
  summarise(avg_delay = mean(Delay, na.rm = TRUE))

ggplot(average_delay, aes(x = vehicle, y = avg_delay)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Average Delay Time by Vehicle Type", x = "Vehicle Type", y = "Average Delay (Minutes)") +
  theme_minimal()

```

```{r ave-delay-by-incident}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: Average Amount of Delay Time by Incident Type
#| fig-align: center

data_incident_avg_delay <- data %>%
  group_by(Incident) %>%
  summarise(avg_delay = mean(Delay, na.rm = TRUE), .groups = 'drop') %>%
  mutate(Incident = fct_reorder(Incident, avg_delay, .desc = TRUE))  # Reorder by avg_delay

# Create the bar chart
ggplot(data_incident_avg_delay, aes(x = Incident, y = avg_delay)) +
  geom_bar(stat = "identity", fill = "steelblue") +  # Set a single color for all bars
  labs(title = "Average Delay Time by Incident Type", 
       x = "Incident Type", y = "Average Delay (Minutes)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10))

data_incident_by_delay <- data %>%
  group_by(Incident, vehicle) %>%
  summarise(incident_count = n(), avg_delay = mean(Delay, na.rm = TRUE), .groups = 'drop') %>%
  group_by(Incident) %>%
  mutate(percentage = (incident_count / sum(incident_count)) * 100) %>%
  mutate(Incident = fct_reorder(Incident, avg_delay, .desc = TRUE))

```

```{r ave-delay-bus}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: Average Amount of Delay Time for Buses by, Incident Type
#| fig-align: center

# Prepare the data: calculate the average delay time for each incident type and vehicle
data_incident_avg_delay <- data %>%
  group_by(vehicle, Incident) %>%
  summarise(avg_delay = mean(Delay, na.rm = TRUE), .groups = 'drop')

# Function to create a bar chart for a specific vehicle
create_bar_chart <- function(vehicle_type) {
  # Filter data for the specific vehicle type and reorder incidents by average delay
  vehicle_data <- data_incident_avg_delay %>%
    filter(vehicle == vehicle_type) %>%
    mutate(Incident = fct_reorder(Incident, avg_delay, .desc = TRUE))  # Reorder by avg_delay

  # Create the bar chart with uniform color and spread the x-axis labels
  ggplot(vehicle_data, aes(x = Incident, y = avg_delay)) +
    geom_bar(stat = "identity", fill = "steelblue") +  # Set a single color for all bars
    labs(title = paste("Average Delay by Incident Type for", vehicle_type),
         x = "Incident Type", y = "Average Delay (Minutes)") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10))  # Rotate labels for readability
}

# Generate the separate bar charts for each vehicle type
bus_bar_chart <- create_bar_chart("Bus")
bus_bar_chart

```

```{r ave-delay-streetcar}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: Average Amount of Delay Time for Streetcars, by Incident Type
#| fig-align: center
streetcar_bar_chart <- create_bar_chart("Streetcar")
streetcar_bar_chart
```

```{r ave-delay-subway}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: Average Amount of Delay Time for Subways, by Incident Type
#| fig-align: center

subway_bar_chart <- create_bar_chart("Subway")
subway_bar_chart

```

```{r ave-delay-days}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: Average Amount of Delay Time by Day
#| fig-align: center
# Step 1: Extract the day of the week from the Date column and calculate average delay time by day of the week
data_avg_delay_by_day <- data %>%
  mutate(day_of_week = wday(Date, label = TRUE, abbr = FALSE)) %>%  # Extract day of the week from Date
  group_by(day_of_week) %>%
  summarise(avg_delay = mean(Delay, na.rm = TRUE), .groups = 'drop')

# Step 2: Create the bar chart for average delay by day of the week
ggplot(data_avg_delay_by_day, aes(x = day_of_week, y = avg_delay)) +
  geom_bar(stat = "identity", fill = "steelblue") +  # Vertical bars for average delay
  labs(title = "Average Delay Time by Day of the Week", 
       x = "Day of the Week", y = "Average Delay Time (Minutes)") +
  theme_minimal()

```

```{r ave-delay-months}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: Average Amount of Delay Time by Month
#| fig-align: center

# Step 1: Extract the month from the Date column and calculate average delay time by month
data_avg_delay_by_month <- data %>%
  mutate(month = month(Date, label = TRUE, abbr = FALSE)) %>%  # Extract month from Date and get full month names
  group_by(month) %>%
  summarise(avg_delay = mean(Delay, na.rm = TRUE), .groups = 'drop')

# Step 2: Reorder the months to display them in chronological order
data_avg_delay_by_month <- data_avg_delay_by_month %>%
  mutate(month = factor(month, levels = month.name))  # Ensure months are ordered chronologically

# Step 3: Create the bar chart with switched axes
ggplot(data_avg_delay_by_month, aes(x = month, y = avg_delay)) +
  geom_bar(stat = "identity", fill = "steelblue") +  # Vertical bars with average delay
  labs(title = "Average Delay Time by Month", 
       x = "Month", y = "Average Delay Time (Minutes)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10))  # Rotate x-axis labels for readability
```

```{r incidents-over-time}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: Number of Incidents Over Time
#| fig-align: center

data_inc_num <- data %>%
  group_by(month = floor_date(Date, "month"), vehicle) %>%
  summarise(incident_count = n())

# Create the line graph
ggplot(data_inc_num, aes(x = month, y = incident_count, color = vehicle, group = vehicle)) +
  geom_line(size = 1) +
  labs(title = "Number of Incidents Over Time by Vehicle", x = "Time", y = "Number of Incidents", color = "Vehicle Type") +
  theme_minimal()

```

```{r delay-over-time}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: Delay Time from 2014 to 2024
#| fig-align: center

data_monthly_delay <- data %>%
  group_by(month = floor_date(Date, "month"), vehicle) %>%
  summarise(average_delay = mean(Delay, na.rm = TRUE), .groups = 'drop')

View(data_monthly_delay)

# Create the line graph for monthly average delay times
ggplot(data_monthly_delay, aes(x = month, y = average_delay, color = vehicle, group = vehicle)) +
  geom_line(size = 1) +
  labs(title = "Average Delay Time Over Time by Vehicle (Monthly)", 
       x = "Time (Month)", y = "Average Delay Time (Minutes)", color = "Vehicle Type") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

# Discussion {#sec-discussion}

## First discussion point {#sec-first-point}

## Second discussion point

## Third discussion point

## Weaknesses and next steps

\newpage

# References